---
title: "Voting round 2"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_download: true
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(message=FALSE,
                      warning=FALSE)
```

# Dependencies

```{r}

library(tidyverse)
library(vote)
library(writexl)

```

# Wrangle data

NB the original data has to be manually wrangled a little to create a combined first row that has the scale and item. find and replace the quesion text, then create an excel formula to paste together the heading and item.

```{r}

dat_all <- read_csv("../data/round 2 - delphi/Delphi round2 cleaned manual.csv") |>
  janitor::clean_names()|>
  rename_with(~ str_replace_all(.x, "_\\d+", "_")) 

dat_design_and_exp <- dat_all |>
  select(starts_with("design_and_exp_"))|>
  select(!ends_with("optional")) |>
  rename_with(~ str_remove_all(.x, "design_and_exp_"))

dat_respsonse_and_gov <- dat_all |>
  select(starts_with("respsonse_and_gov_"))|>
  select(!ends_with("optional")) |>
  rename_with(~ str_remove_all(.x, "respsonse_and_gov_"))

dat_research_and_data <- dat_all |>
  select(starts_with("research_and_data_"))|>
  select(!ends_with("optional")) |>
  rename_with(~ str_remove_all(.x, "research_and_data_"))


dat_comments <- dat_all |>
  select(ends_with("optional")) |>
  pivot_longer(cols = everything(), 
               names_to = "section", 
               values_to = "comment") |>
  arrange(section) |>
  drop_na()

#write_xlsx(dat_comments, "data_processed_comments.xlsx")

```

# Run Single Transferrable Vote 

proportional ranked-choice, multi-winner 

## design and exp

```{r}

res_design_and_exp <- stv(
  votes           = dat_design_and_exp,
  nseats          = 5,
  invalid.partial = TRUE,   # allow partial ballots
  equal.ranking   = FALSE,  # no ties within a ballot
  quota.hare      = FALSE,  # Droop quota (set TRUE for Hare)
  verbose         = FALSE,
  quiet           = TRUE
)

# winners
res_design_and_exp$elected

# # other stuff
# summary(res_design_and_exp)
# plot(res_design_and_exp)
# image(res_design_and_exp, legend=TRUE)

```

## respsonse and gov

```{r}

res_respsonse_and_gov <- stv(
  votes           = dat_respsonse_and_gov,
  nseats          = 5,
  invalid.partial = TRUE,   # allow partial ballots
  equal.ranking   = FALSE,  # no ties within a ballot
  quota.hare      = FALSE,  # Droop quota (set TRUE for Hare)
  verbose         = FALSE,
  quiet           = TRUE
)

# winners
res_respsonse_and_gov$elected

# # other stuff
# summary(res_respsonse_and_gov)
# plot(res_respsonse_and_gov)
# image(res_respsonse_and_gov, legend=TRUE)

```

## research and data

```{r}

res_research_and_data <- stv(
  votes           = dat_research_and_data,
  nseats          = 5,
  invalid.partial = TRUE,   # allow partial ballots
  equal.ranking   = FALSE,  # no ties within a ballot
  quota.hare      = FALSE,  # Droop quota (set TRUE for Hare)
  verbose         = FALSE,
  quiet           = TRUE
)

# winners
res_research_and_data$elected

# # other stuff
# summary(res_research_and_data)
# plot(res_research_and_data)
# image(res_research_and_data, legend=TRUE)

```


