---
title: "Voting round 2"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_download: true
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(message=FALSE,
                      warning=FALSE)
```

# Dependencies

```{r}

library(tidyverse)
library(vote)
library(writexl)

```

# Wrangle data

NB the original data has to be manually wrangled a little to create a combined first row that has the scale and item. find and replace the quesion text, then create an excel formula to paste together the heading and item.

```{r}

dat_all <- read_csv("../data/round 2 - delphi/Delphi round2 cleaned manual.csv") |>
  janitor::clean_names()|>
  rename_with(~ str_replace_all(.x, "_\\d+", "_")) 

dat_design_and_exp <- dat_all |>
  select(starts_with("design_and_exp_"))|>
  select(!ends_with("optional")) |>
  rename_with(~ str_remove_all(.x, "design_and_exp_"))

dat_respsonse_and_gov <- dat_all |>
  select(starts_with("respsonse_and_gov_"))|>
  select(!ends_with("optional")) |>
  rename_with(~ str_remove_all(.x, "respsonse_and_gov_"))

dat_research_and_data <- dat_all |>
  select(starts_with("research_and_data_"))|>
  select(!ends_with("optional")) |>
  rename_with(~ str_remove_all(.x, "research_and_data_"))


dat_comments <- dat_all |>
  select(ends_with("optional")) |>
  pivot_longer(cols = everything(), 
               names_to = "section", 
               values_to = "comment") |>
  arrange(section) |>
  drop_na()

#write_xlsx(dat_comments, "data_processed_comments.xlsx")

```

# Run Single Transferrable Vote 

proportional ranked-choice, multi-winner 

## design and exp

```{r}

res_design_and_exp <- stv(
  votes           = dat_design_and_exp,
  nseats          = 5,
  invalid.partial = TRUE,   # allow partial ballots
  equal.ranking   = FALSE,  # no ties within a ballot
  quota.hare      = FALSE,  # Droop quota (set TRUE for Hare)
  verbose         = FALSE,
  quiet           = TRUE
)

# winners
res_design_and_exp$elected

# # other stuff
# summary(res_design_and_exp)
# plot(res_design_and_exp)
# image(res_design_and_exp, legend=TRUE)

```

## respsonse and gov

```{r}

res_respsonse_and_gov <- stv(
  votes           = dat_respsonse_and_gov,
  nseats          = 5,
  invalid.partial = TRUE,   # allow partial ballots
  equal.ranking   = FALSE,  # no ties within a ballot
  quota.hare      = FALSE,  # Droop quota (set TRUE for Hare)
  verbose         = FALSE,
  quiet           = TRUE
)

# winners
res_respsonse_and_gov$elected

# # other stuff
# summary(res_respsonse_and_gov)
# plot(res_respsonse_and_gov)
# image(res_respsonse_and_gov, legend=TRUE)

```

## research and data

```{r}

res_research_and_data <- stv(
  votes           = dat_research_and_data,
  nseats          = 5,
  invalid.partial = TRUE,   # allow partial ballots
  equal.ranking   = FALSE,  # no ties within a ballot
  quota.hare      = FALSE,  # Droop quota (set TRUE for Hare)
  verbose         = FALSE,
  quiet           = TRUE
)

# winners
res_research_and_data$elected

# # other stuff
# summary(res_research_and_data)
# plot(res_research_and_data)
# image(res_research_and_data, legend=TRUE)

```

# Table 2 (Round 1 priorities and their Round 2 rankings)

Table 2 lists **all** priorities identified in Round 1 with their Round 2 results. For
the **five highest-ranked priorities in each domain** we report the **STV election
order** and the **number of ballots on which the priority was ranked**.
**First-preference counts** are given descriptively, to indicate initial support; they
should **not** be read as determining the final STV order, which incorporates
transferred preferences (see the round-by-round detail below). Priorities carried into
the **World Café** are marked.

Two things worth noting:

* Round 2 used complete drag-and-drop rankings, so every priority was ranked on every
  valid ballot in its domain — the "ballots ranked" column therefore equals the
  domain's valid-ballot count (53 / 52 / 51) for all priorities.
* The World Café discussed the **top two priorities per domain**, taken verbatim from
  the facilitator scripts in `/materials`. In *Governance* and *Research* these are the
  STV top two; in *Design* the World Café used *Harms & benefits* and *Long-term
  effects* (STV #2 and #3) rather than *Dark patterns* (STV #1) — the mismatch is real
  and is flagged here rather than hidden.

```{r}

# The wrangling above ran clean_names(), which drops the human-readable item labels.
# Re-read with the original headers so Table 2 shows readable priority names.
cm_labels <- read_csv("../data/round 2 - delphi/Delphi round2 cleaned manual.csv",
                      show_col_types = FALSE, name_repair = "minimal")

item_cols  <- function(pfx) grep(paste0("^", pfx, "_[0-9]+"), names(cm_labels), value = TRUE)
disp_label <- function(cols)
  str_squish(sub("^(design and exp|respsonse and gov|research and data)_[0-9]+", "", cols))

# Priorities discussed at the World Café = the top two per domain, verbatim from the
# facilitator scripts (materials/Facilitator script - GD&PE / R&G / R&D .docx).
worldcafe_priorities <- c(
  "Harms & benefits",                "Long-term effects",              # Game design
  "Self-regulation & co-regulation", "Child safety & protection",      # Governance
  "Industry-academic data sharing",  "Open data & transparency"        # Research
)

domain_defs <- tibble::tribble(
  ~domain,                             ~prefix,
  "Game design & player experience",   "design and exp",
  "Responsibility & governance",       "respsonse and gov",
  "Research & data",                   "research and data"
)

build_table2_domain <- function(domain, prefix) {
  cols <- item_cols(prefix)
  M    <- sapply(cm_labels[cols], function(x) suppressWarnings(as.numeric(x)))
  d    <- as.data.frame(M[rowSums(!is.na(M)) > 0, , drop = FALSE])
  colnames(d) <- cols                                   # keep original names for STV
  res  <- stv(d, nseats = 5, invalid.partial = TRUE, equal.ranking = FALSE,
              quota.hare = FALSE, verbose = FALSE, quiet = TRUE)
  tibble(
    domain         = domain,
    priority       = disp_label(cols),
    stv_order      = match(cols, res$elected),          # 1..5 for elected, NA otherwise
    ballots_ranked = as.integer(colSums(!is.na(M))),    # ballots ranking the item
    first_pref     = as.integer(colSums(M == 1, na.rm = TRUE)),  # descriptive only
    world_cafe     = disp_label(cols) %in% worldcafe_priorities
  )
}

table2 <- purrr::pmap_dfr(domain_defs, build_table2_domain) |>
  group_by(domain) |>
  # elected five first, in STV election order; remaining priorities by first-pref
  arrange(ifelse(is.na(stv_order), 99L, stv_order), desc(first_pref), .by_group = TRUE) |>
  ungroup()

# manuscript-facing display version
table2_display <- table2 |>
  transmute(
    Domain                   = domain,
    Priority                 = priority,
    `STV election order`     = ifelse(is.na(stv_order), "—", as.character(stv_order)),
    `Ballots ranked (n)`     = ballots_ranked,
    `First-preference count` = first_pref,
    `World Cafe`             = ifelse(world_cafe, "●", "")
  )

knitr::kable(table2_display, caption = "Table 2. Round 1 priorities and Round 2 STV rankings.")

write_xlsx(table2, "../data/outputs/table2_priorities.xlsx")
write_csv(table2,  "../data/outputs/table2_priorities.csv")

```

## Round-by-round detail (elimination order + tallies)

`summary()` prints the full count-by-count tally for each domain, including the order
in which priorities were elected or eliminated.

```{r}

summary(res_design_and_exp)
summary(res_respsonse_and_gov)
summary(res_research_and_data)

```


