---
title: "Round 2 sector diagnostics (exploratory / descriptive)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_download: true
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

# Purpose and scope

A reviewer asked whether different priorities were driven more by different
**sectors**, particularly **academia vs industry**. This file is an **exploratory,
descriptive diagnostic** written to establish what the data can *legitimately*
support before deciding whether to report anything.

It is deliberately kept **separate** from the primary analyses:

* it does **not** modify the pre-registered/primary STV analysis in
  `3_delphi_voting.Rmd`;
* it runs **no** inferential tests (no chi-square, p-values, or confidence
  intervals). The sector sub-groups were not sampled for hypothesis testing and
  this comparison was not pre-registered, so only counts, percentages and rank
  summaries are produced.

# Data source and exact variables used

* **File:** `../data/round 2 - delphi/Delphi round2 cleaned manual.csv`
  (the *same* file used by the STV analysis in `3_delphi_voting.Rmd`, so
  denominators match).
* **Sector:** column `sector` (the multi-select professional-sector item) and
  `sector_other` (free text for "Other").
* **Rankings** (values are rank positions, `1` = most preferred; each *completed*
  domain ballot is a full strict ranking of that domain's items):
  * `design and exp_1` … `design and exp_13`
  * `respsonse and gov_1` … `respsonse and gov_10` (spelling as in the raw export)
  * `research and data_1` … `research and data_7`

The `raw` Qualtrics export (`../data/round 2 - delphi/Delphi raw2.csv`) is used in
one chunk only, to reconcile response counts.

```{r}
library(tidyverse)
library(writexl)

data_path <- "../data/round 2 - delphi/Delphi round2 cleaned manual.csv"
raw_path  <- "../data/round 2 - delphi/Delphi raw2.csv"
out_dir   <- "../data/outputs"
```

# Load

```{r}
cm <- read_csv(data_path, show_col_types = FALSE, name_repair = "minimal") |>
  mutate(rid = row_number())

# ranking columns (the comment columns have no "_<digit>" and are excluded by the regex)
rank_cols <- grep("^(design and exp|respsonse and gov|research and data)_[0-9]+",
                  names(cm), value = TRUE)

# map each ranking column -> domain + human item label
item_map <- tibble(col = rank_cols) |>
  tidyr::extract(col, c("domain_raw", "num", "label"),
          "^(design and exp|respsonse and gov|research and data)_([0-9]+)(.*)$",
          remove = FALSE) |>
  mutate(
    domain = recode(domain_raw,
                    "design and exp"    = "Design & experience",
                    "respsonse and gov" = "Response & governance",
                    "research and data" = "Research & data"),
    label  = str_squish(label),
    num    = as.integer(num)
  )
```

# Sector coding

The survey item text is: *"Which of the following best describes your primary
professional sector? (Select the one that best reflects your role in this space)"*.
Despite the "select the one" wording, the item was exported as a **multiple-select**
field, so a respondent *can* have more than one sector.

**Important parsing note:** the export concatenates the *full* option labels, and
the Industry label itself contains commas
(`"Industry professional (e.g., developer, designer, platform representative,
in-house counsel)"`). Splitting on `,` (or the `;` used in `2_delphi_round2.Rmd`)
therefore mangles sector. We instead detect each canonical option as a **substring**.

```{r}
opts <- c(
  Academia        = "Academic researcher",
  Industry        = "Industry professional",
  `Civil society` = "Civil society or nonprofit representative",
  `Policy/gov`    = "Policy or government advisor",
  Funder          = "Research funder or philanthropic representative",
  Other           = "Other (please specify)"
)

detect_sectors <- function(cell) {
  if (is.na(cell) || cell == "") return(character(0))
  names(opts)[vapply(opts, function(o) str_detect(cell, fixed(o)), logical(1))]
}

cm <- cm |>
  mutate(
    sector_set = lapply(sector, detect_sectors),
    n_sectors  = lengths(sector_set),
    # single-sector ("primary") coding: NA when 0 or >1 sectors selected
    sector_primary = vapply(sector_set, function(x) if (length(x) == 1) x else NA_character_, character(1))
  )

# how many sectors selected
cm |> count(n_sectors) |> knitr::kable(caption = "Number of sectors selected per respondent")
```

# Denominators & reconciliation (n = 59 vs 60 vs 53/52/51)

```{r}
# --- valid ballots per domain in the analysis file (should match STV: 53/52/51) ---
dom_cols <- function(pfx) grep(paste0("^", pfx, "_[0-9]+"), names(cm), value = TRUE)
is_valid <- function(cols) rowSums(!is.na(cm[cols]) & cm[cols] != "") > 0

cm <- cm |>
  mutate(
    valid_design   = is_valid(dom_cols("design and exp")),
    valid_response = is_valid(dom_cols("respsonse and gov")),
    valid_research = is_valid(dom_cols("research and data")),
    any_valid      = valid_design | valid_response | valid_research
  )

recon_cleaned <- tibble(
  quantity = c("Records in analysis file",
               "  with a sector recorded",
               "  with >=1 valid domain ballot",
               "  blank ballots (no rankings at all)",
               "Valid ballots - Design & experience",
               "Valid ballots - Response & governance",
               "Valid ballots - Research & data"),
  n = c(nrow(cm),
        sum(cm$n_sectors > 0),
        sum(cm$any_valid),
        sum(!cm$any_valid),
        sum(cm$valid_design),
        sum(cm$valid_response),
        sum(cm$valid_research))
)
knitr::kable(recon_cleaned, caption = "Analysis-file (cleaned) denominators")

# --- raw Qualtrics export, to explain the n = 59 the manuscript reports ---
raw2 <- read_csv(raw_path, show_col_types = FALSE, name_repair = "minimal") |>
  slice(3:n())  # drop the 2 Qualtrics meta rows (question text + ImportId)

recon_raw <- tibble(
  quantity = c("Raw response rows (meta rows removed)",
               "  with a sector recorded  ->  matches manuscript n",
               "  Finished (Progress = 100)",
               "  with >=1 ranking present"),
  n = c(nrow(raw2),
        sum(!is.na(raw2$sector) & raw2$sector != ""),
        sum(raw2$Finished == "TRUE"),
        sum(rowSums(!is.na(raw2[grep("^(design and exp|respsonse and gov|research and data)_[0-9]+$", names(raw2))]) &
                      raw2[grep("^(design and exp|respsonse and gov|research and data)_[0-9]+$", names(raw2))] != "") > 0))
)
knitr::kable(recon_raw, caption = "Raw-export denominators")
```

**Reading of the discrepancy (not silently resolved):**

* The **raw export has 61** usable response rows. Two have no sector, leaving **59
  with a sector** — this is the most likely source of the manuscript's *n = 59*.
* The **analysis (voting) file has 60** rows. It contains one *fewer* row than the
  raw export; the missing row is an **Industry respondent who submitted no rankings**
  (raw Industry = 16, analysis-file Industry = 15). Why that single blank record was
  dropped during the manual cleaning is **not documented in code** and cannot be
  verified here.
* Only **53 / 52 / 51** rows carry a valid ballot in each domain; these are the STV
  denominators. Seven of the 60 records are **blank ballots** (a sector, sometimes,
  but no rankings).

So *59*, *60* and *53/52/51* count **different things** (respondents-with-sector,
spreadsheet rows, and valid ballots per domain) and are not in conflict once
separated.

# Long-format ranks

```{r}
long <- cm |>
  select(rid, sector_set, all_of(rank_cols)) |>
  pivot_longer(all_of(rank_cols), names_to = "col", values_to = "rank") |>
  mutate(rank = suppressWarnings(as.numeric(rank))) |>
  left_join(item_map |> select(col, domain, label), by = "col") |>
  filter(!is.na(rank))

# explode to one row per (respondent x sector) so multi-sector respondents count
# in each sector they selected
long_sec <- long |> unnest_longer(sector_set, values_to = "sector")

# valid ballots per sector x domain (the correct denominators for %)
valid_by <- long_sec |>
  distinct(rid, sector, domain) |>
  count(sector, domain, name = "n_valid")
```

# (3) Sector sample sizes

```{r}
# multi-select membership (a respondent contributes to every sector they picked)
sector_membership <- tibble(sector = unlist(cm$sector_set)) |>
  count(sector, name = "n_selecting", sort = TRUE)

# mutually-exclusive "primary" (single-sector respondents only)
sector_primary_counts <- cm |> count(sector_primary, name = "n_primary")

# valid ballots available per sector, by domain
valid_wide <- valid_by |> pivot_wider(names_from = domain, values_from = n_valid, values_fill = 0)

sector_Ns <- sector_membership |>
  left_join(cm |> filter(n_sectors == 1) |> count(sector_primary, name = "n_single"),
            by = c("sector" = "sector_primary")) |>
  left_join(valid_wide, by = "sector") |>
  mutate(interpretable = if_else(n_selecting >= 10, "yes", "TOO SMALL - unstable"))

knitr::kable(sector_Ns, caption = "Sector Ns (multi-select membership, single-sector count, and valid ballots by domain)")
knitr::kable(sector_primary_counts, caption = "Mutually-exclusive primary sector (single-sector respondents; NA = 0 or >1 sectors)")
```

Only **Academia (~26–27)** and **Industry (~15)** have samples large enough for even
descriptive comparison. **Policy/gov (6), Civil society (5), Other (5), Funder (1)**
are too small — a single respondent moves their percentages by 15–100 points — and
are reported for completeness only.

# (4) Priority preferences by sector — descriptive

```{r}
# ---- first preference: item ranked 1 ----
first_pref <- long_sec |>
  filter(rank == 1) |>
  count(sector, domain, label, name = "n_first") |>
  left_join(valid_by, by = c("sector", "domain")) |>
  mutate(pct_first = round(100 * n_first / n_valid, 1))

# ---- top-5 inclusion: item ranked <= 5 ----
top5 <- long_sec |>
  filter(rank <= 5) |>
  count(sector, domain, label, name = "n_top5") |>
  left_join(valid_by, by = c("sector", "domain")) |>
  mutate(pct_top5 = round(100 * n_top5 / n_valid, 1))

# ---- rank distribution summary (mean/median rank among those who ranked it) ----
rank_dist <- long_sec |>
  group_by(sector, domain, label) |>
  summarise(n_ranked = n(), mean_rank = round(mean(rank), 2),
            median_rank = median(rank), .groups = "drop")

# tidy full tables (all sectors kept; small-sector flag included)
reliable <- function(s) if_else(s %in% c("Academia", "Industry"), "", " (small N - unstable)")

first_pref_out <- first_pref |>
  arrange(domain, sector, desc(pct_first)) |>
  mutate(note = reliable(sector))
top5_out <- top5 |>
  arrange(domain, sector, desc(pct_top5)) |>
  mutate(note = reliable(sector))
```

## Academia vs Industry — TOP-5 inclusion (%)

```{r}
ai_top5 <- top5 |>
  filter(sector %in% c("Academia", "Industry")) |>
  select(domain, label, sector, pct_top5) |>
  pivot_wider(names_from = sector, values_from = pct_top5, values_fill = 0) |>
  mutate(diff = round(Academia - Industry, 1)) |>
  arrange(domain, desc(Academia))
knitr::kable(ai_top5, caption = "Top-5 inclusion %: Academia (n~26) vs Industry (n~14). diff = Academia - Industry")
```

## Academia vs Industry — FIRST preference (%)

```{r}
ai_first <- first_pref |>
  filter(sector %in% c("Academia", "Industry")) |>
  select(domain, label, sector, n_first, pct_first) |>
  pivot_wider(names_from = sector, values_from = c(n_first, pct_first), values_fill = 0) |>
  arrange(domain, desc(pct_first_Academia))
knitr::kable(ai_first, caption = "First-preference counts/% : Academia vs Industry")
```

# (5) Exploratory sector-specific STV (Academia & Industry only)

Run **only** for Academia (~26 ballots) and Industry (~14 ballots), alongside the
full sample for comparison. **This is an exploratory sensitivity check, not a
result.** Industry in particular is small: with a Droop quota, 5 seats from ~14
ballots means each "winner" is backed by only ~3 first-preferences, so its ordering
is fragile. STV is **not** run for Policy/Civil society/Other/Funder (too small).

```{r}
library(vote)

short_lab <- function(cols) str_squish(sub("^(design and exp|respsonse and gov|research and data)_[0-9]+", "", cols))
in_sector <- function(s) vapply(cm$sector_set, function(x) s %in% x, logical(1))

run_stv_safe <- function(dat) {
  dat <- as.data.frame(lapply(dat, function(x) suppressWarnings(as.numeric(x))))
  dat <- dat[rowSums(!is.na(dat)) > 0, , drop = FALSE]      # valid ballots only
  suppressWarnings(stv(dat, nseats = 5, invalid.partial = TRUE, equal.ranking = FALSE,
                       quota.hare = FALSE, verbose = FALSE, quiet = TRUE))
}

domains <- list(
  `Design & experience`   = dom_cols("design and exp"),
  `Response & governance` = dom_cols("respsonse and gov"),
  `Research & data`       = dom_cols("research and data")
)

stv_compare <- map_dfr(names(domains), function(dn) {
  cols <- domains[[dn]]
  grp <- list(All = cm, Academia = cm[in_sector("Academia"), ], Industry = cm[in_sector("Industry"), ])
  map_dfr(names(grp), function(g) {
    d <- grp[[g]][cols]; colnames(d) <- cols
    n_valid <- sum(rowSums(!is.na(sapply(d, as.numeric))) > 0)
    res <- run_stv_safe(d)
    tibble(domain = dn, group = g, n_ballots = n_valid,
           elected_in_order = paste(short_lab(res$elected), collapse = " > "))
  })
})
knitr::kable(stv_compare, caption = "Exploratory STV elected sets (order of election) by group")
```

# (7) Write outputs

```{r}
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

# compact academia-vs-industry supplementary table (top-5 % + diff, per domain)
academia_vs_industry <- ai_top5

write_xlsx(list(sector_Ns = sector_Ns,
                sector_primary = sector_primary_counts),
           file.path(out_dir, "sector_Ns.xlsx"))
write_xlsx(first_pref_out, file.path(out_dir, "sector_first_pref_by_priority.xlsx"))
write_xlsx(top5_out,       file.path(out_dir, "sector_top5_by_priority.xlsx"))
write_xlsx(list(top5_diff = academia_vs_industry,
                first_pref = ai_first,
                stv_compare = stv_compare),
           file.path(out_dir, "academia_vs_industry_supp.xlsx"))

# CSV mirrors
write_csv(sector_Ns,      file.path(out_dir, "sector_Ns.csv"))
write_csv(first_pref_out, file.path(out_dir, "sector_first_pref_by_priority.csv"))
write_csv(top5_out,       file.path(out_dir, "sector_top5_by_priority.csv"))
write_csv(rank_dist,      file.path(out_dir, "sector_rank_distribution.csv"))
write_csv(academia_vs_industry, file.path(out_dir, "academia_vs_industry_top5.csv"))
```

# Session info

```{r}
sessionInfo()
```
